### 2015-12-22 ismsmemo/literature.tex

% $Id$
% Author:	Daniel Bosk <daniel.bosk@miun.se>
Du ska inför skrivningen av detta PM ha läst dokumenten
\begin{itemize}
  \item \emph{Introduktion till metodstödet}~\cite{MSB2011itm},
  \item \emph{Säkra ledningens engagemang}~\cite{MSB2011sle}, och
  \item \emph{Projektplanering}~\cite{MSB2011p}.
\end{itemize}