### 2017-03-16 msbframework/abstract.tex
Adds WIP abstract for msbframework (WIP #4)

This lecture covers the first part of MSB's framework~\cite{%
  MSB2011itm,MSB2011sle,MSB2011p,MSB2011v,MSB2011r%
}, i.e.~ISO 27001.
This part covers how to initialise the work with security in an organisation, 
i.e.~how to set up an Information Security Management System (ISMS).
We will talk about the most important steps in this process.

This lecture covers the remaining part of MSB's material~\cite{%
	MSB2011gap,MSB2011vs,MSB2011us,MSB2011upo,%
	MSB2011pg,MSB2011koa,MSB2011i,MSB2011o,MSB2011g,%
	MSB2011lg,MSB2011ulo,MSB2011kf,MSB2011fa%
}.
This part of the material treats how to run an ISMS\@.
The largest part is the gap analysis, i.e.~finding the gap between the security 
practices in the organisation and the practices recommended by ISO 27000.
The main point of this part is not something done once and never again, an ISMS 
is a continuous process.