### 2017-03-22 ismsmemo/abstract.tex
Translates literature of assignments (WIP #3)
Adds the translated literature as abstract.tex. Later to be replaced by
a thorough abstract (see #4, #5, #17).

Before writing this memo, you should have read the following:
\begin{itemize}
  \item \citetitle{MSB2011itm}~\cite{MSB2011itm},
  \item \citetitle{MSB2011sle}~\cite{MSB2011sle}, and
  \item \citetitle{MSB2011p}~\cite{MSB2011p}.
\end{itemize}