### 2018-03-22 ismsmemo/abstract.tex

This assignment is purely for assessment of the following \acp{ILO}, you should 
be able to
\begin{itemize}
  \item \emph{understand} the importance of working a strategically with 
    information security and what \iac{ISMS} is.
  \item \emph{value} which content is more or less important in \iac{ISMS}.
  \item \emph{describe} which success factors are important when implementing 
    \iac{ISMS}.
\end{itemize}

The literature consists of the following documents:
\begin{itemize}
  \item \citetitle{MSB2011itm}~\cite{MSB2011itm},
  \item \citetitle{MSB2011sle}~\cite{MSB2011sle}, and
  \item \citetitle{MSB2011p}~\cite{MSB2011p}.
\end{itemize}