### 2018-03-22 ismsmemo/literature.tex

% $Id$
% Author:	Daniel Bosk <daniel.bosk@miun.se>, Lennart Franked
% <lennart.franked@miun.se>

Before starting this PM, you should have read the following documents:

\begin{itemize}
  \item \emph{Introduktion till metodstödet}~\cite{MSB2011itm},
  \item \emph{Säkra ledningens engagemang}~\cite{MSB2011sle}, och
  \item \emph{Projektplanering}~\cite{MSB2011p}.
\end{itemize}

Alternatively
\begin{itemize}
  \item \cite[Chapter 3]{iso27000}

\end{itemize}