### 2019-03-05 msbframework/msbforts/abstract.tex

% $Id$
This lecture covers the remaining part of MSB's material~\cite{%
	MSB2011gap,MSB2011vs,MSB2011us,MSB2011upo,%
	MSB2011pg,MSB2011koa,MSB2011i,MSB2011o,MSB2011g,%
	MSB2011lg,MSB2011ulo,MSB2011kf,MSB2011fa%
}.
This part of the material treats how to run an ISMS\@.
The largest part is the gap analysis, i.e.~finding the gap between the security 
practices in the organisation and the practices recommended by ISO 27000.
The main point of this part is not something done once and never again, an ISMS 
is a continuous process.