### 2019-03-25 gapproject/abstract.tex
Adds an abstract for gapproject

In this project we will explore information security management in 
organizations further.
In particular, it will provide some practical experience.

After this project you should be able
\begin{itemize}
  \item to \emph{analyse} threats to information security and to \emph{suggest} 
    suitable protection mechanisms against these.
  \item to \emph{apply} MSB's framework to analyse, evaluate and give 
    suggestions of improvements for the information security in an 
    organization.
\end{itemize}

The work is based on the material by MSB:
\citetitle{MSB2011gap}~\cite{MSB2011gap} med bilaga~\cite{MSB2011gb}.