### 2022-03-21 ismsmemo/ismsmemo.tex
Updated instructions

\documentclass[a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[swedish]{babel}
\usepackage{csquotes}
\usepackage[hyphens]{url}
\usepackage{hyperref}
\usepackage[swedish]{cleveref}
\usepackage[single]{acro}
\usepackage[natbib,style=numeric-comp,maxbibnames=99,sorting=none]{biblatex}
\addbibresource{ismsmemo.bib}

\usepackage{verbatim}
\let\solution\comment%

\pagestyle{plain}

\DeclareAcronym{isms}
{
  short = ISMS,
  short-indefinite = an,
  long = information security management system,
  long-indefinite = an,
}

% XXX translate to english, add enisa literature

\begin{document}
\title{PM\@: Information Security Management System}
\author{%
  Carina Bengtsson and Lennart Franked
}
\date{\today}

\maketitle

\section{Introduction}
\label{sec:introduction}

This assignments focus on \acp{isms}. The purpose of \iac{isms} is to manage
an organizations work within the field of information security.

\section{Aim}
\label{sec:aim}

This assignment aims to:
\begin{itemize}
  \input{aims.tex}
\end{itemize}

\section{Reading assignment}

\input{literature.tex}

\section{Task}\label{Work}

This assignment contains four sub tasks, One introduction task covering
\acp{isms}, including Swedish statutes, success stories and how to get a
committed leadership.

\subsection{Information Security Management System}
Shortly describe a management system is, and more specifically what \iac{isms} 
is. Your description should also cover how \iac{isms} is
structured.

\subsection{Legal requirements according to MSBFS 2020:6}

In \iac{isms}, there are certain work tasks defined for the organisation.
The Swedish Civil Contingencies Agency (MSB) have defined in their statutes
MSBFS 2020:6~\cite{MSBFS2020:6}, which is binding for governmental agencies, 
stated that the governmental agencies should:
\begin{enumerate}
  \item Ensure that there is an information security policy where the
        management's objectives and focus on
        the information security work are shown,
  \item Clarify the management and the rest of the organizations
        responsibilities, including the person or persons appointed to lead and coordinate
        information security work, and give these positions the
        powers needed,
  \item ensure that the information security work is assigned the necessary resources,
  \item establish the necessary internal rules, working methods and support 
  \item ensure that the content of the organisation's internal rules, working methods and support is
    evaluated and, if necessary, adapted.

\end{enumerate}
Further more, in 14§ that

\begin{enumerate}\setcounter{enumi}{5}
  \item the management should keep itself up to date with the work with
    information security, and at least once a year do a follow-up and evaluate
    the information security work.
\end{enumerate}

Shortly reflect around the key points 1, 2 and 3, \emph{then} select one of the key
points 4, 5 or 6.

If you are using the English course material, refer to the key-points a-i in
\cite[chap. 4.2.1]{iso27000}

Write down why you think each of the requirement is important, or not as
important to incorporate in an organisation.

Note that the organisation does not necessarily need to be a governmental
agency.

\subsection{Success factors}

To implement \iac{isms} in an organisation is a large project. Therefore
the material mentions some success factors that are critical to achieve a
successful implementation \cite[4.6]{iso27000}. Give a short summary about
these success factors.

\subsection{Commitment from the Management}
If you where put in the position to try to convince the management in a company
to implement a \iac{isms}. How would you motivate them?

\section{Examination}
\label{sec:examination}

This assignment will be examinated through a written PM that will be handed in
as a PDF-document.
The PM \emph{must} must cover the tasks from \cref{Work}:

\begin{enumerate}
  \item Information Security Management System;
  \item Requirements according to MSBFS 2020:6, points 1--3 and one of the
    following key points 4--6. Or if you are using the english material,
    keypoints a-c and one the points d-i in \cite[chap. 4.2.1]{iso27000}.
  \item Success factors; and
  \item Commitment from the Management
\end{enumerate}

Your PM should be written with an academic language in either english or
swedish. It must contain correct references.

\printbibliography{}
\end{document}