### 2022-03-21 ismsmemo/literature.tex

% $Id$
% Author:	Daniel Bosk <daniel.bosk@miun.se>, Lennart Franked
% <lennart.franked@miun.se>

Before starting this PM, you should have read the following documents:

\begin{itemize}
  \item \emph{Myndigheten för samhällsskydd och beredskaps föreskrifter om informationssäkerhet för
      statliga myndigheter}~\cite{MSBFS2020:6}
  \item \emph{Information technology – Security techniques – Information
security management systems – Overview and vocabulary (ISO/IEC 27000:2018)}~\cite[chap. 4]{iso27000}
(Use as a support document only, if you don't understand Swedish, use this as the main reference)
  \item \emph{Introduktion till metodstödet}~\cite{MSB2018iom}, och
  \item \emph{Ledningens engagemang avgörande}~\cite{MSB2018sle}
\end{itemize}

Alternatively
\begin{itemize}
  \item ISO 27000 standard~\cite[Chapter 4]{iso27000}
\end{itemize}